### 2023-07-18 Article_PySF.tex
myzhu update
update

In this work, an open-source program implementing WFAT in the integral representation is developed that can easily and quickly calculate the molecular structure factor of common molecules. In our implementation, named the \texttt{PyStructureFactor}, the ionizing molecular orbital is expanded as a linear combination of GTOs at the HF level of theory calculated by the \texttt{PySCF} package \cite{sun_pyscf_2018, sun_recent_2020}. The integrals are evaluated numerically using a three-dimensional grid-based method, which exhibits much better numerical stability and is readily applicable to large molecules, compared to the standard analytical integration procedure. Our present implementation of \texttt{PyStructureFactor} features \cyan{flexible control of accuracy and} efficient computation towards the calculation of the molecular structure factor.

@@

The next topic is concerned with the calculation of the HF potential operator acting on the ionizing orbital, i.e., $\cyan{V}(\bn{r}) \psi_0(\bn{r})$. In the practical application, the molecular orbitals are usually obtained by the self-consistent-field (SCF) approach, which is commonly implemented in standard quantum chemistry packages. The molecular orbitals are usually expanded into linear combinations of Gaussian-type orbitals (GTOs):
\begin{equation}
    \psi_i(\bn{r})=\sum_\alpha C_{i \alpha} \chi_\alpha(\bn{r}-\bn{R}_\alpha),
\end{equation}
where $\chi_\alpha(\bn{r}-\bn{R}_\alpha)$ denotes a GTO of index $\alpha$ centering at $\bn{R}_\alpha$, and the coefficients $C_{i \alpha}$ are obtained through the SCF approach.
In this way we obtained $V_{\rm{d}}(\bn{r})$ and $V_{\rm{ex}}(\bn{r})\psi_0(\bn{r})$ in Eqs.~(\ref{eq:Vd}) and (\ref{eq:VexPsi0}) expressed in coefficients $C_{i \alpha}$ for practical calculations. Defining the integral
\begin{equation}
    J_{\alpha\beta}(\bn{r}) = \int \frac{\chi_\alpha(\bn{r}'-\bn{R}_\alpha) \chi_\beta(\bn{r}'-\bn{R}_\beta)}{\abs{\bn{r}-\bn{r}'}} \dd \bn{r}'
\end{equation}
which can be calculated analytically for GTOs,
Eqs.~(\ref{eq:Vd}) and (\ref{eq:VexPsi0}) can be written as
\begin{align}
     V_{\rm{d}}(\bn{r}) &= \sum_{i=1}^{N} \sum_{\alpha,\beta} C_{i\alpha} C_{i\beta} J_{\alpha\beta}(\bn{r}), \\
     \cyan{V}_{\rm{ex}}(\bn{r})\psi_0(\bn{r}) &= - \sum_{i=1}^{N} \braket{\sigma_i | \sigma_0} \sum_{\alpha,\beta,\gamma} C_{i\alpha} C_{i\beta} C_{0\gamma} J_{\beta\gamma}(\bn{r}) \chi_\alpha(\bn{r}-\bn{R}_\alpha).
\end{align}

@@

@@

                                      on the $(\beta,\gamma)$ orientation grid, whose shape is \cout{initialized from} \cyan{given by the} \texttt{orient\_grid\_size}. \\

@@

Now we present convergence and reliability tests of the program in Figs.~\ref{fig:convtest_N2} and \ref{fig:convtest_CO} with N$_2$ and CO molecules, respectively. To this end, we obtain their structure factors as a function of $\beta$ and compare the results to reference data extracted from the literature \cite{saito_structure_2015} which is obtained using the tail representation of WFAT. In both figures, the pc-4 basis is adopted in the calculations, and \cyan{the cut-off limit of the index $l$ in the summation of Eq.~(\ref{eq:StructFactorExpInt}) , i.e., the } $l_{\rm{max}}$, is set to the default value of 10. The ``grid\_level'' labels denote the input parameter \texttt{``atom\_grid\_level''} of the program. Here, grid levels 1, 3, 5, and 7 correspond to radial and angular sizes (40, 194), (75, 302), (105, 770), and (135, 1202) for each spherical grid around the nuclei, respectively. It is clear from the figures that for \texttt{``atom\_grid\_level''} larger than 3, the obtained results are hardly distinguishable from each other even after enlargement, which reveals a stable and fast convergence of the present program. In addition, the converged result is close to the reference data. This illustrates that the calculation of the molecular structure factor is able to achieve accurate results under default parameters. Minor discrepancy from the reference data results from the difference in the tail and integral presentations of WFAT and the choice of basis.

>>>>

In this work, an open-source program implementing WFAT in the integral representation is developed that can easily and quickly calculate the molecular structure factor of common molecules. In our implementation, named the \texttt{PyStructureFactor}, the ionizing molecular orbital is expanded as a linear combination of GTOs at the HF \cyan{or complete-active-space self-consistent field (CASSCF)} level of theory calculated by the \texttt{PySCF} package \cite{sun_pyscf_2018, sun_recent_2020}. The integrals are evaluated numerically using a three-dimensional grid-based method, which exhibits much better numerical stability and is readily applicable to large molecules, compared to the standard analytical integration procedure. Our present implementation of \texttt{PyStructureFactor} features \cyan{flexible control of accuracy and} efficient computation towards the calculation of the molecular structure factor.

@@

The next topic is concerned with the calculation of the HF potential operator acting on the ionizing orbital, i.e., $\cyan{V}(\bn{r}) \psi_0(\bn{r})$. In the practical application, the molecular orbitals are usually obtained by the self-consistent-field (SCF) approach, which is commonly implemented in standard quantum chemistry packages. The molecular orbitals are usually expanded into linear combinations of Gaussian-type orbitals (GTOs):
\begin{equation}
    \psi_i(\bn{r})=\sum_\alpha C_{i \alpha} \chi_\alpha(\bn{r}-\bn{R}_\alpha),
\end{equation}
where $\chi_\alpha(\bn{r}-\bn{R}_\alpha)$ denotes a GTO of index $\alpha$ centering at $\bn{R}_\alpha$, and the coefficients $C_{i \alpha}$ are obtained through the SCF approach.
In this way we obtain $V_{\rm{d}}(\bn{r})$ and $V_{\rm{ex}}(\bn{r})\psi_0(\bn{r})$ in Eqs.~(\ref{eq:Vd}) and (\ref{eq:VexPsi0}) expressed in coefficients $C_{i \alpha}$ for practical calculations. Defining the integral
\begin{equation}
    J_{\alpha\beta}(\bn{r}) = \int \frac{\chi_\alpha(\bn{r}'-\bn{R}_\alpha) \chi_\beta(\bn{r}'-\bn{R}_\beta)}{\abs{\bn{r}-\bn{r}'}} \dd \bn{r}'
\end{equation}
which can be calculated analytically for GTOs,
Eqs.~(\ref{eq:Vd}) and (\ref{eq:VexPsi0}) can be written as
\begin{align}
     V_{\rm{d}}(\bn{r}) &= \sum_{i=1}^{N} \sum_{\alpha,\beta} C_{i\alpha} C_{i\beta} J_{\alpha\beta}(\bn{r}), \\
     \cyan{V}_{\rm{ex}}(\bn{r})\psi_0(\bn{r}) &= - \sum_{i=1}^{N} \braket{\sigma_i | \sigma_0} \sum_{\alpha,\beta,\gamma} C_{i\alpha} C_{i\beta} C_{0\gamma} J_{\beta\gamma}(\bn{r}) \chi_\alpha(\bn{r}-\bn{R}_\alpha).
\end{align}

@@

        \cyan{\texttt{casscf\_conf}}& \cyan{Configuration of CASSCF calculation consisting of \texttt{(n\_active\_orb, n\_active\_elec)}.
                                      Specifying \texttt{None} (by default) indicates employing the primitive HF method instead of the more accurate CASSCF method.}                          & \cyan{\texttt{None}}\\

@@

                                      on the $(\beta,\gamma)$ orientation grid, whose shape is \cout{initialized from} \cyan{given by} \texttt{orient\_grid\_size}. \\

@@

Now we present convergence and reliability tests of the program in Figs.~\ref{fig:convtest_N2} and \ref{fig:convtest_CO} with N$_2$ and CO molecules, respectively. To this end, we obtain their structure factors as a function of $\beta$ and compare the results to reference data extracted from the literature \cite{saito_structure_2015} which is obtained using the tail representation of WFAT. In both figures, the pc-4 basis is adopted in the calculations, and \cyan{the cut-off limit of the index $l$ in the summation of Eq.~(\ref{eq:StructFactorExpInt}) , i.e.,} $l_{\rm{max}}$, is set to the default value of 10. The ``grid\_level'' labels denote the input parameter \texttt{``atom\_grid\_level''} of the program. Here, grid levels 1, 3, 5, and 7 correspond to radial and angular sizes (40, 194), (75, 302), (105, 770), and (135, 1202) for each spherical grid around the nuclei, respectively. It is clear from the figures that for \texttt{``atom\_grid\_level''} larger than 3, the obtained results are hardly distinguishable from each other even after enlargement, which reveals a stable and fast convergence of the present program. In addition, the converged result is close to the reference data. This illustrates that the calculation of the molecular structure factor is able to achieve accurate results under default parameters. Minor discrepancy from the reference data results from the difference in the tail and integral presentations of WFAT and the choice of basis.
